---
title: |
  | "PH700A Week 4"
subtitle: |
  | Lab Assignment 4
  | PH 700A, Spring 2025
author: "FirstName LastName"
format: html
editor: source
---

## Data Information

This data originated from PhysioNet.

Please view the documentation for the data contents at: <https://mimic.mit.edu/docs/iv/modules/ed/>

## Data Import and Package Load
```{r}
#| title: data import
#| include: false

library(data.table)
library(tidyverse)

df.ed <- fread("https://github.com/core-academics/sdsu.ph700a/raw/refs/heads/main/data/mimic4ed/edstays.csv.gz")

df.dx <- fread("https://github.com/core-academics/sdsu.ph700a/raw/refs/heads/main/data/mimic4ed/diagnosis.csv.gz")

df.triage <- fread("https://github.com/core-academics/sdsu.ph700a/raw/refs/heads/main/data/mimic4ed/triage.csv.gz")

df.dxwide <- df.dx %>% 
  pivot_wider(
    id_cols = c(subject_id, stay_id),
    names_from = seq_num,
    values_from = c(icd_code, icd_version, icd_title)
  )

df.ed <- df.ed %>% mutate(hlos = outtime-intime)

df.ed <- df.ed %>% mutate(across(c("gender", "race"), as.factor))

df.triage <- df.triage %>% mutate(fever = if_else(temperature >= 100.4, 1, 0))
```

## Scenario

We are tasked by the hospital to identify factors associated with long hospital length of stay (`df.ed$hlos`) among patients _admitted to the hospital_. Variables of interest include:

-   Method of Arrival to the Hospital: `df.ed$arrival_transport == "ADMITTED"`

-   Hypotensive Shock at Admission defined as a systolic blood pressure < 90 mmHg: `df.triage$sbp < 100`

-   High reported pain score > 8: `df.triage$pain > 8`

-   Tachycardia defined as heart rate > 100: `df.triage$heartrate > 100`

-   Any acute or chronic kidney disease: `df.dx$icd_title ~ "kidney"`

## Objective

Evaluate the specified risk factors for their relationship with hospital length of stay.

## Data Preparation

### Isolate Hospital Admissions

```{r}
df <- subset(df.ed, disposition == "ADMITTED")
```

### Define Kidney Disease
```{r}
#| label: scanning icd_title for kidney

df.dx$icd_title <- tolower(df.dx$icd_title)

df.kidney <- df.dx %>%
  mutate(kidneydisease = if_else(grepl("kidney", icd_title), 1, 0))

df.kidney <- subset(df.kidney, kidneydisease == 1)

df.kidney <- df.kidney %>% select(stay_id, kidneydisease)

df <- left_join(df, df.kidney, by = "stay_id")

df$kidneydisease <- replace_na(df$kidneydisease,0)

# object cleanup
rm(df.kidney)
```

Explanation:

The `grepl` is an extension of the `grep` function and stands for, "global regular expression print" and is used for scanning text for character patterns.

Here, `grepl` is used to look for the word, "kidney" in the variable `icd_title`. The `if_else` command assigns a 1 to the new variable `kidneydisease` if the pattern is found, and a 0 if it is not found.

### Generate Measure for Hypotensive Shock

### Generate Measure for High Pain Score

### Generate Measure for Tachycardia

### Combine variables of interest in a unified data frame

### Evaluate if any of the above risk factors are associated with hospital length of stay

