---
title: |
  | "PH700A Week 3"
subtitle: |
  | Lab Assignment 3
  | PH 700A, Spring 2025
author: "FirstName LastName"
format: html
editor: source
---

## Data Information

This data originated from PhysioNet.

Please view the documentation for the data contents at: <https://mimic.mit.edu/docs/iv/modules/ed/>

## Data Import and Package Load
```{r}
#| title: data import
#| include: false

library(data.table)
library(tidyverse)

df.ed <- fread("https://github.com/core-academics/sdsu.ph700a/raw/refs/heads/main/data/mimic4ed/edstays.csv.gz")

df.dx <- fread("https://github.com/core-academics/sdsu.ph700a/raw/refs/heads/main/data/mimic4ed/diagnosis.csv.gz")

df.triage <- fread("https://github.com/core-academics/sdsu.ph700a/raw/refs/heads/main/data/mimic4ed/triage.csv.gz")
```

## Data Frame Contents
```{r}
glimpse(df.ed)

glimpse(df.dx)

glimpse(df.triage)
```

#  Objectives

## Create a new column to capture the total time spent in the emergency department
```{r}

```

## Encode gender and race as factors
```{r}

```

## Identify records that were admitted with a fever of 100.4 or greater
```{r}

```

## Reshape the diagnosis table to wide
```{r}
df.dxwide <- df.dx %>% 
  pivot_wider(
    id_cols = c(subject_id, stay_id),
    names_from = seq_num,
    values_from = c(icd_code, icd_version, icd_title)
  )
```

## Identify only records where "INTRACEREBRAL HEMORRHAGE" was reported as the FIRST diagnosis
```{r}

```


## Isolate only records with "INTRACEREBRAL HEMORRHAGE" in a new object
```{r}

```

## Consolidate the new factors you created in a unified data frame
```{r}

```

## Are there any patients without any diagnosis codes? If so, what is their subject_id?
```{r}

```

